---
title: "Platelets gene analysis"
output:
  html_document:
    df_print: paged
  pdf_document: default
---


Data were processed with the nf-core RNA-seq pipeline https://github.com/nf-core/rnaseq which includes STAR and salmon. Further analysis was performed using the R package DEseq2.

**Differential expression analysis overview:**

[RPs vs MPs - CCS only](#RPsMPs_CCS)
- design = ~ RPs.MPs + patient (DE between RPs/MPs with paired patients)
- data: 38 CCS samples
- DE statistics for all genes: results/ccs_rp_vs_mp_diff_gene_counts.xlsx
- only significant genes: results/ccs_rp_vs_mp_sign_diff_gene_counts.xlsx
- gene set enrichment: results/ccs_rp_vs_mp_sign_gene_set_enrichment.xlsx


```{r echo=FALSE, warning=FALSE, message=FALSE}
setwd("D:/Uni/Hiwi_ExBioLab/Rechts-der-Isar/Material/analyses_in_R/CCSonly_withoutSample6_filtered_restructured")
source("source/create_plots.R")
library(DESeq2)
library(data.table)
library(pheatmap)
library(ggplot2)
library(ggpubr)
library(RColorBrewer)
library(ggrepel)
library(viridis)
library(xlsx)
library(dplyr)
library(webshot)
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

plot_width = 8.2
plot_height = 5.8
plot_res = 300
version <- "_noSample6_filtered"
preplot_title <- "Before filtering"
post_filter_title <- "Genes expressed with at least 10 counts in at least 20% of samples"
post_normalization_title <- "After normalization with DESeq2"
write_tables = FALSE
```



```{r echo=FALSE, warning=FALSE, message=FALSE}
source("../source/sample_annotation.R")
metadata_ccs <- get_sample_annotation(picked_disease = "CCS", remove_samples = c("6_RPs", "6_MPs"))
if(write_tables){
  write.table(metadata_ccs, paste0("../data/sample_annotation/processed_metadata_ccs_only", version, ".csv"), sep = ";", quote = FALSE, row.names = FALSE)
}
#metadata_ccs <- read.csv(paste0("../data/gene_counts/processed_metadata_ccs_only", version, ".csv"), sep = ";")

plot <- myAgeDistrubutionPlot(metadata_ccs)
ggsave(plot = plot, path ='plots', filename = paste0("age_distribution", version, ".png"),
       width = plot_width, height = plot_height, dpi = plot_res, units = "in", device='png')
ggsave(plot = plot, path ='plots/pdf', filename = paste0("age_distribution", version, ".pdf"),
       width = plot_width, height = plot_height, dpi = plot_res, units = "in", device='pdf')
plot
```


```{r echo=FALSE, warning=FALSE, message=FALSE}
source("../source/process_input.R")
counts_data <- get_raw_gene_counts(metadata_ccs)
counts_ccs <- counts_data$counts
id2symbol <- counts_data$id2symbol
rm(counts_data)
if(write_tables){
  write.table(counts_ccs, paste0("./processed_raw_counts_ccs_only", version, ".csv"), sep = ";", quote = FALSE, row.names = FALSE)
}
#counts_ccs <- read.csv(paste0("../data/gene_counts/processed_raw_counts_ccs_only", version, ".csv"), sep = ";")
#colnames(counts_ccs) <- gsub("X", "", colnames(counts_ccs))

pre_plot <- myCountAnalysisPlot(counts_ccs, "gene", preplot_title)
pre_plot
```


```{r, echo=FALSE, message= FALSE, warning=FALSE}
# pre-filtering of rows having less than 10 counts in total as suggested by DESeq2
expressed_samples <- data.frame(apply(counts_ccs, 1, function(x) sum(x != "0")))
colnames(expressed_samples) <- "count_expressed_samples"
expressed_samples$gene <- rownames(expressed_samples)
counts_ccs <- counts_ccs[rownames(counts_ccs) %in% rownames(expressed_samples[expressed_samples$count_expressed_samples >= ncol(counts_ccs)*0.2 ,]) & rowSums(counts_ccs) >= 10,]
rm(expressed_samples)

postfilterplot <- myCountAnalysisPlot(counts_ccs, "gene", post_filter_title)
postfilterplot
```


```{r echo=FALSE, warning=FALSE, message=FALSE}
dds_ccs <- DESeqDataSetFromMatrix(countData = counts_ccs,
                                             colData = metadata_ccs,
                                             design = ~ patient + RPs_MPs)
dds_ccs <- DESeq(dds_ccs)
dds_ccs_counts <- counts(dds_ccs, norm = TRUE)
dds_ccs_results <- data.frame(results(dds_ccs, contrast=c("RPs_MPs","RPs","MPs")))

postnormalizationplot <- myCountAnalysisPlot(as.data.frame(dds_ccs_counts), "gene",
                                             post_normalization_title)
postnormalizationplot


plot <- ggarrange(pre_plot, postfilterplot, postnormalizationplot, labels = c("A", "B", "C"), ncol = 3, nrow = 1)
rm(pre_plot, postfilterplot, postnormalizationplot)
ggsave(plot = plot, path ='plots', filename = paste0("CCS_only_expressionAnalysis", version, ".png"),
       width = plot_width * 3, height = plot_height, dpi = plot_res, units = "in", device='png')
ggsave(plot = plot, path ='plots/pdf', filename = paste0("CCS_only_expressionAnalysis", version, ".pdf"),
       width = plot_width * 3, height = plot_height, dpi = plot_res, units = "in", device='pdf')
plot
```


```{r echo=FALSE, warning=FALSE, message=FALSE}
remove_padj_na_recursive <- FALSE
remove_padj_na <- TRUE

if(remove_padj_na_recursive){
  genes_with_na_padj <- rownames(dds_ccs_results[is.na(dds_ccs_results$padj),])
  while(length(genes_with_na_padj) != 0){
    dds_ccs <- dds_ccs[which(!rownames(dds_ccs) %in% genes_with_na_padj), ]
    dds_ccs_counts <- counts(dds_ccs, norm = TRUE)
    dds_ccs_results <- data.frame(results(dds_ccs, contrast=c("RPs_MPs","RPs","MPs")))
    genes_with_na_padj <- rownames(dds_ccs_results[is.na(dds_ccs_results$padj),])
  }
  
  postnormalizationfilterplot <- myCountAnalysisPlot(as.data.frame(dds_ccs_counts), "gene", 
                                                     "After removing genes with padj is NA labeled by DESeq2")
  ggsave(plot = postnormalizationfilterplot, path ='plots', 
         filename = paste0("CCS_only_expressionAnalysis_postnormalizationfilterplot", version, ".png"),
         width = plot_width, height = plot_height, dpi = plot_res, units = "in", device='png')
  ggsave(plot = postnormalizationfilterplot, path ='plots/pdf', 
         filename = paste0("CCS_only_expressionAnalysis_postnormalizationfilterplot", version, ".pdf"),
         width = plot_width, height = plot_height, dpi = plot_res, units = "in", device='pdf')
  postnormalizationfilterplot
} else if(remove_padj_na) {
    genes_with_na_padj <- rownames(dds_ccs_results[is.na(dds_ccs_results$padj),])
    dds_ccs <- dds_ccs[which(!rownames(dds_ccs) %in% genes_with_na_padj), ]
    dds_ccs_counts <- dds_ccs_counts[which(!rownames(dds_ccs_counts) %in% genes_with_na_padj), ]
    dds_ccs_results <- dds_ccs_results[which(!rownames(dds_ccs_results) %in% genes_with_na_padj), ]
} else {
  cat("No manual removal of genes with padj = NA done")
}
```


```{r echo=FALSE, warning=FALSE, message=FALSE}
postfilter_biotypedata <- myBiotypeBarplot(counts_ccs, paste0("CCS only - ", post_filter_title))
biotype_barplot_postfilter <- postfilter_biotypedata$biotype_plot
biotype_table_postfilter <- postfilter_biotypedata$biotype_table
rm(postfilter_biotypedata)
ggsave(plot = biotype_barplot_postfilter, path ='plots', 
       filename = paste0("CCS_only_biotypeAnalysis_postfilterplot", version, ".png"),
       width = plot_width, height = plot_height, dpi = plot_res, units = "in", device='png')
ggsave(plot = biotype_barplot_postfilter, path ='plots/pdf', 
       filename = paste0("CCS_only_biotypeAnalysis_postfilterplot", version, ".pdf"),
       width = plot_width, height = plot_height, dpi = plot_res, units = "in", device='pdf')
gtsave(biotype_table_postfilter, path = ('plots'), 
       filename = paste0("CCS_only_biotypeAnalysis_postfiltertable", version, ".png"))
gtsave(biotype_table_postfilter, path = ('plots/pdf'), 
       filename = paste0("CCS_only_biotypeAnalysis_postfiltertable", version, ".pdf"), zoom = 1)
rm(biotype_barplot_postfilter, biotype_table_postfilter)


postnormalization_biotypedata <- myBiotypeBarplot(as.data.frame(dds_ccs_counts), paste0("CCS only - ", post_normalization_title))
biotype_barplot_postnormalization <- postnormalization_biotypedata$biotype_plot
biotype_table_postnormalization <- postnormalization_biotypedata$biotype_table
rm(postfilter_biotypedata)
ggsave(plot = biotype_barplot_postnormalization, path ='plots', 
       filename = paste0("CCS_only_biotypeAnalysis_postnormalizationplot", version, ".png"),
       width = plot_width, height = plot_height, dpi = plot_res, units = "in", device='png')
ggsave(plot = biotype_barplot_postnormalization, path ='plots/pdf', 
       filename = paste0("CCS_only_biotypeAnalysis_postnormalizationplot", version, ".pdf"),
       width = plot_width, height = plot_height, dpi = plot_res, units = "in", device='pdf')
gtsave(biotype_table_postnormalization, path = ('plots'), 
       filename = paste0("CCS_only_biotypeAnalysis_postnormalizantiontable", version, ".png"))
gtsave(biotype_table_postnormalization, path = ('plots/pdf'), 
       filename = paste0("CCS_only_biotypeAnalysis_postnormalizationtable", version, ".pdf"), zoom = 1)
```



```{r, warning=FALSE, message= FALSE}
threshold <- sort(rowVars(dds_ccs_counts), decreasing = T)[150]
dds_ccs_top = dds_ccs_counts[which(rowVars(dds_ccs_counts) >= threshold),]
plot <- pheatmap(dds_ccs_top, 
                 fontsize_col = 7,
                 color = colorRampPalette(c("blue", "red"))(50),
                 border_color = NA,
                 annotation_colors = list(`Platelet type` = c(RPs = cbbPalette[1], MPs = cbbPalette[4])),
                 annotation_col = metadata_ccs %>% dplyr::select(`Platelet type` = RPs_MPs),
                 show_rownames = FALSE,
                 show_colnames = TRUE,
                 scale="row")
rm(threshold, dds_ccs_top)

ggsave(plot = plot, path ='plots', filename = paste0("CCS_only_heatmap_top150", version, ".png"),
       width = plot_width, height = plot_height, dpi = plot_res, units = "in", device='png')
ggsave(plot = plot, path ='plots/pdf', filename = paste0("CCS_only_heatmap_top150", version, ".pdf"),
       width = plot_width, height = plot_height, dpi = plot_res, units = "in", device='pdf')
plot
```



```{r warning=FALSE, message= FALSE}
dds_ccs_results <- as.data.frame(dds_ccs_results)
dds_ccs_results <- merge(dds_ccs_results, id2symbol, by.x = 0, by.y = 'ensg')
rownames(dds_ccs_results) <- dds_ccs_results$Row.names
dds_ccs_results <- dds_ccs_results[,c("hgnc_symbol", "baseMean", "log2FoldChange", "lfcSE", "stat", "pvalue", "padj", "gene_biotype")]
dds_ccs_results <- dds_ccs_results[order(dds_ccs_results[,'padj']),]

write.xlsx2(dds_ccs_results, paste0("results/ccs_rp_vs_mp_diff_gene_counts", version, ".xlsx"))
DT::datatable(dds_ccs_results)
```


```{r, warning=FALSE, message= FALSE}
fdr <- 0.1
fc <- 1

plot_nolim <- myMAplot(dds_ccs_results, fdr = fdr, fc = fc)
ggsave(plot = plot_nolim, path ='plots', filename = paste0("CCS_only_ma_noLim", version, ".png"),
       width = plot_width, height = plot_height, dpi = plot_res, units = "in", device='png')
ggsave(plot = plot_nolim, path ='plots/pdf', filename = paste0("CCS_only_ma_noLim", version, ".pdf"),
       width = plot_width, height = plot_height, dpi = plot_res, units = "in", device='pdf')

plot_lim2 <- myMAplot_lim(dds_ccs_results, fdr = fdr, fc = fc, ylim = c(-2, 2))
ggsave(plot = plot_lim2, path ='plots', filename = paste0("CCS_only_ma_lim2", version, ".png"),
       width = plot_width, height = plot_height, dpi = plot_res, units = "in", device='png')
ggsave(plot = plot_lim2, path ='plots/pdf', filename = paste0("CCS_only_ma_lim2", version, ".pdf"),
       width = plot_width, height = plot_height, dpi = plot_res, units = "in", device='pdf')

plot_lim4 <- myMAplot_lim(dds_ccs_results, fdr = fdr, fc = fc, ylim = c(-4, 4))
ggsave(plot = plot_lim4, path ='plots', filename = paste0("CCS_only_ma_lim4", version, ".png"),
       width = plot_width, height = plot_height, dpi = plot_res, units = "in", device='png')
ggsave(plot = plot_lim4, path ='plots/pdf', filename = paste0("CCS_only_ma_lim4", version, ".pdf"),
       width = plot_width, height = plot_height, dpi = plot_res, units = "in", device='pdf')

plot_lim7 <- myMAplot_lim(dds_ccs_results, fdr = fdr, fc = fc, ylim = c(-7, 7))
ggsave(plot = plot_lim7, path ='plots', filename = paste0("CCS_only_ma_lim7", version, ".png"),
       width = plot_width, height = plot_height, dpi = plot_res, units = "in", device='png')
ggsave(plot = plot_lim7, path ='plots/pdf', filename = paste0("CCS_only_ma_lim7", version, ".pdf"),
       width = plot_width, height = plot_height, dpi = plot_res, units = "in", device='pdf')

plot <- ggarrange(plot_nolim, plot_lim2, plot_lim4, plot_lim7, labels = c("no lim", "lim 2", "lim 4", "lim 7"), ncol = 2, nrow = 2)
rm(plot_nolim, plot_lim2, plot_lim4, plot_lim7)
plot
```


Next, we show a heatmap focusing only on the top 20, 50 and 100 differentially expressed genes having the smallest p-values from the above list.

```{r, warning=FALSE, message=FALSE}
top_20 <- dds_ccs_counts[rownames(dds_ccs_results)[1:20],]
rownames(top_20) <- dds_ccs_results$hgnc_symbol[1:20]
plot <- pheatmap(top_20,
                 fontsize_col = 7,
                 fontsize_row = 10,
                 color = colorRampPalette(c("blue", "red"))(50), 
                 border_color = NA,
                 annotation_colors = list(`Platelet type` = c(RPs = cbbPalette[1], MPs = cbbPalette[4])),
                 annotation_col = metadata_ccs %>% dplyr::select(`Platelet type` = RPs_MPs),
                 show_rownames = TRUE, 
                 show_colnames = TRUE, 
                 scale="row")
rm(top_20)

ggsave(plot = plot, path ='plots', filename = paste0("CCS_only_heatmap_top20", version, ".png"),
       width = plot_width, height = plot_height, dpi = plot_res, units = "in", device='png')
ggsave(plot = plot, path ='plots/pdf', filename = paste0("CCS_only_heatmap_top20", version, ".pdf"),
       width = plot_width, height = plot_height, dpi = plot_res, units = "in", device='pdf')
plot
```


```{r, warning=FALSE, message=FALSE}
top_50 <- dds_ccs_counts[rownames(dds_ccs_results)[1:50],]
rownames(top_50) <- dds_ccs_results$hgnc_symbol[1:50]
plot <- pheatmap(top_50,
                 fontsize_col = 7,
                 fontsize_row = 7,
                 color = colorRampPalette(c("blue", "red"))(50), 
                 border_color = NA,
                 annotation_colors = list(`Platelet type` = c(RPs = cbbPalette[1], MPs = cbbPalette[4])),
                 annotation_col = metadata_ccs %>% dplyr::select(`Platelet type` = RPs_MPs),
                 show_rownames = TRUE, 
                 show_colnames = TRUE, 
                 scale="row")
rm(top_50)

ggsave(plot = plot, path ='plots', filename = paste0("CCS_only_heatmap_top50", version, ".png"),
       width = plot_width, height = plot_height, dpi = plot_res, units = "in", device='png')
ggsave(plot = plot, path ='plots/pdf', filename = paste0("CCS_only_heatmap_top50", version, ".pdf"),
       width = plot_width, height = plot_height, dpi = plot_res, units = "in", device='pdf')
plot
```


```{r, warning=FALSE, message=FALSE}
top_100 <- dds_ccs_counts[rownames(dds_ccs_results)[1:100],]
rownames(top_100) <- dds_ccs_results$hgnc_symbol[1:100]
plot <- pheatmap(top_100,
                 fontsize_col = 7,
                 fontsize_row = 4,
                 color = colorRampPalette(c("blue", "red"))(50), 
                 border_color = NA,
                 annotation_colors = list(`Platelet type` = c(RPs = cbbPalette[1], MPs = cbbPalette[4])),
                 annotation_col = metadata_ccs %>% dplyr::select(`Platelet type` = RPs_MPs),
                 show_rownames = TRUE, 
                 show_colnames = TRUE, 
                 scale="row")
rm(top_100)

ggsave(plot = plot, path ='plots', filename = paste0("CCS_only_heatmap_top100", version, ".png"),
       width = plot_width, height = plot_height, dpi = plot_res, units = "in", device='png')
ggsave(plot = plot, path ='plots/pdf', filename = paste0("CCS_only_heatmap_top100", version, ".pdf"),
       width = plot_width, height = plot_height, dpi = plot_res, units = "in", device='pdf')
plot
```


Here, the samples are colored by platelet type and labeled by patient.

```{r, warning=FALSE, message= FALSE}
plot <- myPCAplot(vst(dds_ccs), intgroup = "RPs_MPs") + 
          geom_label_repel(aes(label=colnames(dds_ccs)), size = 2, show.legend = FALSE)

ggsave(plot = plot, path ='plots', filename = paste0("CCS_only_pca_RPvsMP", version, ".png"),
       width = plot_width, height = plot_height, dpi = plot_res, units = "in", device='png')
ggsave(plot = plot, path ='plots/pdf', filename = paste0("CCS_only_pca_RPvsMP", version, ".pdf"),
       width = plot_width, height = plot_height, dpi = plot_res, units = "in", device='pdf')
plot
```


In this PCA plot, the samples are colored by patient and labeled by platelet type. The variance is better explained by patient rather than by platelet type.
```{r, warning=FALSE, message= FALSE}
plot <- myPCAplot(vst(dds_ccs), intgroup = c("patient")) + 
          geom_label_repel(aes(label=RPs_MPs), size = 2) + 
          theme(legend.position="none")

ggsave(plot = plot, path ='plots', filename = paste0("CCS_only_pca_patiens", version, ".png"),
       width = plot_width, height = plot_height, dpi = plot_res, units = "in", device='png')
ggsave(plot = plot, path ='plots/pdf', filename = paste0("CCS_only_pca_patiens", version, ".pdf"),
       width = plot_width, height = plot_height, dpi = plot_res, units = "in", device='pdf')
plot
```


## Differentially expressed genes in CCS
Below we list the significant differentially expressed genes between RPs and MPs when matching patients. (abs(log2FC) > 1, padj < 0.1)

```{r, warning=FALSE, message= FALSE}
FCcutoff <- 1
pCutoff <- 0.1

write.table(rownames(dds_ccs_results), paste0("../data/GO_analysis/reference_genes_ccs_only", version, ".txt"),
            sep = ",", quote = FALSE, row.names = FALSE, col.names = FALSE)

upregulated_genes <- dds_ccs_results %>%
  filter(log2FoldChange > FCcutoff, padj < pCutoff) %>%
  arrange(padj)
write.table(rownames(upregulated_genes), 
            paste0("../data/GO_analysis/sign_upregulated_genes_padj", pCutoff, "_ccs_only", version, ".txt"), 
            sep = ",", quote = FALSE, row.names = FALSE, col.names = FALSE)

downregulated_genes <- dds_ccs_results %>%
  filter(log2FoldChange < -FCcutoff, padj < pCutoff) %>%
  arrange(padj)
write.table(rownames(downregulated_genes), 
            paste0("../data/GO_analysis/sign_downregulated_genes_padj", pCutoff, "_ccs_only", version, ".txt"), 
            sep = ",", quote = FALSE, row.names = FALSE, col.names = FALSE)

dds_ccs_sign_results <- dds_ccs_results %>% 
  tibble::rownames_to_column() %>% 
  filter(abs(log2FoldChange) > FCcutoff, padj < pCutoff) %>%
  arrange(padj)
write.xlsx2(dds_ccs_sign_results, paste0("results/ccs_rp_vs_mp_sign_diff_padj_", pCutoff, "_gene_counts", version, ".xlsx"), row.names = F)
DT::datatable(dds_ccs_sign_results)
```


Volcano plot of the significant differences, here filtered for log2 FC +-1 and p-adj < 1e-15.

```{r warning=FALSE, message=FALSE}
FCcutoff <- 1
pCutoff <- 1e-15

upregulated_genes <- dds_ccs_results[dds_ccs_results$log2FoldChange > FCcutoff & dds_ccs_results$padj < pCutoff, ]
upregulated_genes <- upregulated_genes[order(upregulated_genes$padj, decreasing = FALSE), ]
write.table(rownames(upregulated_genes), 
            paste0("../data/GO_analysis/sign_upregulated_genes_padj", pCutoff, "_ccs_only", version, ".txt"), 
            sep = ",", quote = FALSE, row.names = FALSE, col.names = FALSE)

downregulated_genes <- dds_ccs_results[dds_ccs_results$log2FoldChange < -FCcutoff & dds_ccs_results$padj < pCutoff, ]
downregulated_genes <- downregulated_genes[order(downregulated_genes$padj, decreasing = FALSE), ]
write.table(rownames(downregulated_genes), 
            paste0("../data/GO_analysis/sign_downregulated_genes_padj", pCutoff, "_ccs_only", version, ".txt"), 
            sep = ",", quote = FALSE, row.names = FALSE, col.names = FALSE)


labs_of_interest <- c("GP6", "PEAR1", "VWF", "PKN2", "LUC7L3", "PARN", "ORAI2", "STIM1", "TBXA2R",
                                "F2RL3", "P2RX1", "GP5", "ITGA5", "GPVI", "PEAR")

cat(paste0("Number of IDs of interest in current data: ", length(labs_of_interest[labs_of_interest %in% dds_ccs_results$hgnc_symbol]), "\nList of detected IDs of interes: "))
cat(labs_of_interest[labs_of_interest %in% dds_ccs_results$hgnc_symbol])
```


```{r warning=FALSE, message=FALSE}
plot <- myVolcanoPlot(dds_ccs_results,
                      data_type = "gene",
                      pCutoff = pCutoff,
                      FCcutoff = FCcutoff,
                      xlim = c(-6, 6),
                      lab_selection = labs_of_interest)

ggsave(plot = plot, path ='plots', filename = paste0("CCS_only_volcano_geneset", version, ".png"),
       width = plot_width, height = plot_height, dpi = plot_res, units = "in", device='png')
ggsave(plot = plot, path ='plots/pdf', filename = paste0("CCS_only_volcano_geneset", version, ".pdf"),
       width = plot_width, height = plot_height, dpi = plot_res, units = "in", device='pdf')
plot
```


```{r warning=FALSE, message=FALSE}
regulated_labs_of_interest <- labs_of_interest[labs_of_interest %in% c(upregulated_genes$hgnc_symbol, downregulated_genes$hgnc_symbol)]

cat(paste0("Number of regulated IDs of interest in current data: ", length(regulated_labs_of_interest), "\nList of detected IDs of interes: "))
cat(regulated_labs_of_interest)
```


```{r warning=FALSE, message=FALSE}
plot <- myVolcanoPlot(dds_ccs_results,
                      data_type = "gene",
                      pCutoff = pCutoff,
                      FCcutoff = FCcutoff,
                      xlim = c(-6, 6),
                      lab_selection = regulated_labs_of_interest)

ggsave(plot = plot, path ='plots', filename = paste0("CCS_only_volcano_regGeneset", version, ".png"),
       width = plot_width, height = plot_height, dpi = plot_res, units = "in", device='png')
ggsave(plot = plot, path ='plots/pdf', filename = paste0("CCS_only_volcano_regGeneset", version, ".pdf"),
       width = plot_width, height = plot_height, dpi = plot_res, units = "in", device='pdf')
plot
```


```{r warning=FALSE, message=FALSE}
top_10_upregulated_without_labs_of_interest <- as.character(upregulated_genes[!(upregulated_genes$hgnc_symbol %in% labs_of_interest),"hgnc_symbol"])[1:10]
plot <- myVolcanoPlot(dds_ccs_results,
                      data_type = "gene",
                      pCutoff = pCutoff,
                      FCcutoff = FCcutoff,
                      xlim = c(-6, 6),
                      lab_selection = c(regulated_labs_of_interest, top_10_upregulated_without_labs_of_interest))

ggsave(plot = plot, path ='plots', filename = paste0("CCS_only_volcano_regGenset_top10upreg", version, ".png"),
       width = plot_width, height = plot_height, dpi = plot_res, units = "in", device='png')
ggsave(plot = plot, path ='plots/pdf', filename = paste0("CCS_only_volcano_regGenset_top10upreg", version, ".pdf"),
       width = plot_width, height = plot_height, dpi = plot_res, units = "in", device='pdf')
plot
```


```{r warning=FALSE, message=FALSE}
top_10_downregulated_without_labs_of_interest <- as.character(downregulated_genes[!(downregulated_genes$hgnc_symbol %in% labs_of_interest),"hgnc_symbol"])[1:10]

plot <- myVolcanoPlot(dds_ccs_results,
                      data_type = "gene",
                      pCutoff = pCutoff,
                      FCcutoff = FCcutoff,
                      xlim = c(-6, 6),
                      lab_selection = c(regulated_labs_of_interest,
                                        top_10_downregulated_without_labs_of_interest))

ggsave(plot = plot, path ='plots', filename = paste0("CCS_only_volcano_regGenset_top10downreg", version, ".png"),
       width = plot_width, height = plot_height, dpi = plot_res, units = "in", device='png')
ggsave(plot = plot, path ='plots/pdf', filename = paste0("CCS_only_volcano_regGenset_top10downreg", version, ".pdf"),
       width = plot_width, height = plot_height, dpi = plot_res, units = "in", device='pdf')
plot
```


```{r warning=FALSE, message=FALSE}
plot <- myVolcanoPlot(dds_ccs_results,
                      data_type = "gene",
                      pCutoff = pCutoff,
                      FCcutoff = FCcutoff,
                      xlim = c(-6, 6),
                      lab_selection = c(regulated_labs_of_interest, top_10_upregulated_without_labs_of_interest,
                                        top_10_downregulated_without_labs_of_interest))

ggsave(plot = plot, path ='plots', filename = paste0("CCS_only_volcano_regGenset_top10up&downreg", version, ".png"),
       width = plot_width, height = plot_height, dpi = plot_res, units = "in", device='png')
ggsave(plot = plot, path ='plots/pdf', filename = paste0("CCS_only_volcano_regGenset_top10up&downreg", version, ".pdf"),
       width = plot_width, height = plot_height, dpi = plot_res, units = "in", device='pdf')
plot
```


## Gene onthology analysis
The following interactive plot shows the classes of genes that are over-represented in the list of significant differentially expressed genes in CCS samples (RPs vs MPs paired by patient).

```{r warning=FALSE, message=FALSE}
#test
padj <- padj_threshold <- 0.01
count <- count_threshold <- 5
fdr <- fdr_threshold <- 0.2


gos_of_interest <- c("platelet", "hemostasis", "coagulation")
padj_thresholds <- c(0.1, 0.01, 0.05, 1e-15)
count_thresholds <- c(2, 5, 10, 12, 15, 20)
fdr_thresholds <- c(0.01, 0.05, 0.1, 0.2, 0.5, 0.75, 1)

for(padj in padj_thresholds){
  panther_result_up_ref <- fread(paste0("../data/GO_analysis/padj", padj, 
                                    "/Panther_upregulatedGenes_ReferenceBackground_padj", padj, ".txt"))
  panther_result_down_ref <- fread(paste0("../data/GO_analysis/padj", padj,
                                      "/Panther_downregulatedGenes_ReferenceBackground_padj", padj, ".txt"))
  panther_result_up_human <- fread(paste0("../data/GO_analysis/padj", padj, 
                                    "/Panther_upregulatedGenes_HumanBackground_padj", padj, ".txt"))
  panther_result_down_human <- fread(paste0("../data/GO_analysis/padj", padj,
                                      "/Panther_downregulatedGenes_HumanBackground_padj", padj, ".txt"))
  
  
  for(count in count_thresholds){
    for(fdr in fdr_thresholds){
      plot_ref <- myGObarplot(panther_result_up = panther_result_up_ref, 
                              panther_result_down = panther_result_down_ref, 
                              count_threshold = count, 
                              fdr_threshold = fdr,
                              padj_threshold = padj,
                              gos_of_interest = gos_of_interest,
                              background = "Reference genes")
      plot_human <- myGObarplot(panther_result_up = panther_result_up_human, 
                                panther_result_down = panther_result_down_human, 
                                count_threshold = count, 
                                fdr_threshold = fdr,
                                padj_threshold = padj,
                                gos_of_interest = gos_of_interest,
                                background = "All human genes")
      plot <- ggarrange(plot_ref, plot_human, ncol = 1, nrow = 2)
      ggsave(plot = plot, path = paste0("plots/padj", padj), 
             filename = paste0("CCS_only_GOanalysis_padj", padj, "_count", count, "_fdr", fdr, version, ".png"),
             width = plot_width*3, height = plot_height*2, dpi = plot_res, units = "in", device='png')
    }
  }
}

```




# R session
```{r}
sessionInfo()
```
