---
title: "Plateletes miRNA analysis"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

Data were processed with the nf-core small RNA-seq pipeline https://github.com/nf-core/smrnaseq which includes mirtop https://doi.org/10.1093/bioinformatics/btz675 to generate output in mirGFF3 format. Further analysis was performed using the R packages isoMiR, DEseq2 and ggplot2.

**Differential expression analysis overview:**

[RPs vs MPs - CCS only](#RPsMPs_CCS)
- design = ~ patient + RPs.MPs (DE between RPs/MPs with paired patients)
- data: 40 CCS samples
- DE statistics for all genes: results/ccs_rp_vs_mp_diff_mir_counts.xlsx
- only significant genes: results/ccs_rp_vs_mp_sign_diff_mir_counts.xlsx

```{r echo=FALSE, warning=FALSE, message=FALSE}
setwd("D:/Uni/Hiwi_ExBioLab/Rechts-der-Isar/Material/analyses_in_R/CCSonly_withoutSample6_filtered_restructured")
source("../source/create_plots.R")
library(DESeq2)
library(ggplot2)
library(ggrepel)
library(viridis)
library(xlsx)
library(RColorBrewer)
library(dplyr)
library(pheatmap)
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

plot_width = 8.2
plot_height = 5.8
plot_res = 300
version <- "_noSample6_filtered_pCutoff0.05"
preplot_title <- "Before filtering"
post_filter_title <- "miRNAs expressed with at least 10 counts in at least 20% of samples"
post_normalization_title <- "After normalization with DESeq2"

fdr = 0.1
pCutoff = 0.05
FCcutoff = 0.5
```



```{r, echo = FALSE, warning=FALSE, message= FALSE}
source("../source/sample_annotation.R")
metadata_ccs <- get_sample_annotation(picked_disease = "CCS", remove_samples = c("6_RPs", "6_MPs"))


source("../source/process_input.R")
counts_ccs <- get_mirtop_mirna_counts(metadata_ccs)

pre_plot <- myCountAnalysisPlot(counts_ccs, "miRNA", preplot_title)
pre_plot
```



```{r, echo = FALSE, warning=FALSE, message= FALSE}
# pre-filtering of rows having less than 10 counts in total as suggested by DESeq2
expressed_samples <- data.frame(apply(counts_ccs, 1, function(x) sum(x != "0")))
colnames(expressed_samples) <- "count_expressed_samples"
expressed_samples$gene <- rownames(expressed_samples)
counts_ccs <- counts_ccs[rownames(counts_ccs) %in% rownames(expressed_samples[expressed_samples$count_expressed_samples >= ncol(counts_ccs)*0.2,]) & rowSums(counts_ccs) >= 10,]
rm(expressed_samples)

postfilterplot <- myCountAnalysisPlot(counts_ccs, "miRNA", post_filter_title)
postfilterplot
```


```{r echo=FALSE, warning=FALSE, message=FALSE}
dds_ccs <- DESeqDataSetFromMatrix(countData = counts_ccs,
                              colData = metadata_ccs,
                              design = ~ patient + RPs_MPs)
dds_ccs <- DESeq(dds_ccs)
dds_ccs_results <- as.data.frame(results(dds_ccs, contrast=c("RPs_MPs","RPs","MPs")))
dds_ccs_counts <- as.data.frame(counts(dds_ccs, norm = TRUE))
rm(counts_ccs)

postnormalizationplot <- myCountAnalysisPlot(dds_ccs_counts, "miRNA", post_normalization_title)
postnormalizationplot


plot <- ggarrange(pre_plot, postfilterplot, postnormalizationplot, labels = c("A", "B", "C"), ncol = 3, nrow = 1)
ggsave(path ='plots/pdf', filename = paste0("CCS_only_expressionAnalysis", version, ".pdf"),
       width = plot_width * 3, height = plot_height, dpi = plot_res, units = "in", device='pdf')
ggsave(path ='plots/png', filename = paste0("CCS_only_expressionAnalysis", version, ".png"),
       width = plot_width * 3, height = plot_height, dpi = plot_res, units = "in", device='png')
plot
rm(plot, pre_plot, postfilterplot, postnormalizationplot)
```


```{r echo=FALSE, warning=FALSE, message=FALSE}
num_top = 15
plot <- myTopmiRNABarplot(dds_ccs_counts, num_top)
ggsave(path ='plots/pdf', filename = paste0("All_top", num_top, "_most_expressed_mirnas", version, ".pdf"),
       width = plot_width, height = plot_height, dpi = plot_res, units = "in", device='pdf')
ggsave(path ='plots/png', filename = paste0("All_top", num_top, "_most_expressed_mirnas", version, ".png"),
       width = plot_width, height = plot_height, dpi = plot_res, units = "in", device='png')
plot
```


# RPs vs MPs - CCS only {#RPsMPs_CCS}

Finally, we repeat the differential gene expression analysis for CCS patients only, as the sample quality is expected to be higher here.


```{r, warning=FALSE, message= FALSE}
dds_ccs_counts <- as.matrix(dds_ccs_counts)
plot <- pheatmap(dds_ccs_counts[which(rowVars(dds_ccs_counts) > 0.1),], 
                 fontsize_col = 7,
                 color = colorRampPalette(c("navy", "red"))(50),
                 border_color = NA,
                 annotation_colors = list(`Platelet type` = c(RPs = cbbPalette[1], MPs = cbbPalette[4])),
                 annotation_col = metadata_ccs %>% dplyr::select(`Platelet type` = RPs_MPs),
                 show_rownames = FALSE, 
                 show_colnames = TRUE, 
                 scale="row")

png(paste0("plots/CCS_only_heatmap", version, ".png"), width = plot_width, height = plot_height, units = "in", res = plot_res)
plot
invisible(dev.off())
plot
```

```{r, warning=FALSE, message= FALSE}
plot <- myMAplot(dds_ccs_results, fdr = fdr, fc = FCcutoff)

png(paste0("plots/CCS_only_ma", version, ".png"), width = plot_width, height = plot_height, units = "in", res = plot_res)
plot
invisible(dev.off())
plot
```


PCA of CCS patients colored by MPs/RPs and labeled by sample:
```{r, warning=FALSE, message= FALSE}
plot <- myPCAplot(rlog(dds_ccs), intgroup = c("RPs_MPs")) + 
          geom_label_repel(aes(label=sample_name), size = 2, show.legend = FALSE)

png(paste0("plots/CCS_only_pcs_RPvsMP", version, ".png"), width = plot_width, height = plot_height, units = "in", res = plot_res)
plot
invisible(dev.off())
plot
```


PCA of CCS patients colored by patient and labeled by MPs/RPs. The variance seems to be stronger between patients than between MPs/RPs.
```{r, warning=FALSE, message= FALSE}
plot <- myPCAplot(rlog(dds_ccs), intgroup = c("patient")) + 
          geom_label_repel(aes(label=`RPs_MPs`), size = 2) + theme(legend.position="none")

png(paste0("plots/CCS_only_pca_patients", version, ".png"), width = plot_width, height = plot_height, units = "in", res = plot_res)
plot
invisible(dev.off())
plot
```

In the CCS samples, there are no significant differentially expressed miRNAs in RPs vs MPs for |log2FoldChange| > 0.5 and padj < 0.1.
```{r, warning=FALSE, message= FALSE}
write.xlsx2(dds_ccs_results, paste0("results/ccs_rp_vs_mp_diff_mir_counts", version, ".xlsx"))
dds_ccs_sign_results <- dds_ccs_results %>% 
  tibble::rownames_to_column() %>% 
  filter(abs(log2FoldChange) > FCcutoff, padj < pCutoff) %>%
  arrange(padj)
write.xlsx2(dds_ccs_sign_results, paste0("results/ccs_rp_vs_mp_sign_diff_mir_counts", version, ".xlsx"), row.names = F)
DT::datatable(dds_ccs_sign_results)
```

Volcano plot using a p-adjusted cutoff of 0.01 and a log2 FC cutoff of +-0.5.

```{r, warning=FALSE, message= FALSE}
plot <- myVolcanoPlot(dds_ccs_results,
                      data_type = "miRNA",
                      pCutoff = pCutoff,
                      FCcutoff = FCcutoff,
                      xlim = c(-1, 1),
                      ylim = c(0, 3))

png(paste0("plots/CCS_only_volcano", version, ".png"), width = plot_width, height = plot_height, units = "in", res = plot_res)
plot
invisible(dev.off())
plot
```


# R session
```{r}
sessionInfo()
```
